Dependent Object Types

### 2012-08-09 article.tex
typog of wfe

%% Relations
% Subtype 
\newcommand{\sub}{<:}
% Type assignment
\newcommand{\typ}{:}
% reduction
\newcommand{\reduces}{\;\rightarrow\;}
% well-formedness
\newcommand{\wf}{\;\mbox{\textbf{wf}}}
\newcommand{\wfe}{\;\mbox{\textbf{wfe}}}

@@

The last rule, (\textsc{new}), assigns types to object creation
expressions. It is the most complex of DOT's typing rules.  To
type-check an object creation $\new y {T_c \ldefs {\seq{l =
      v}\;\seq{m(x) = t}}} t'$, one verifies first that the type $T_c$
is well-formed (see \figref{dot-wf} for a definition of
well-formedness).  One then determines the set of all declarations
that this type carries, using the expansion relation $\expand$ defined
in \figref{dot-two}.  Every type declaration $\Ldecl L S U$ in
this set must be realizable, i.e.\ its lower bound $S$ must be a
subtype of its upper bound $U$.  Every field declaration $\ldecl l V$
in this set must have a corresponding initializing value of $v$ of
type $V$.  These checks are made in an environment which is extended
by the binding $y: T_c$. In particular this allows field values that
recurse on ``self'' by referring to the bound variable $x$. Similarly,
every method declaration $\mdecl m T W$ must have a corresponding
initializing method definition $m(x) = t$. The parameter type $T$ must
be $\wfe$ (well-formed and expanding; see \figref{dot-wf}), and
the body $t$ must type check to $W$ in an environment extended by the
bindings $y : T_c$ and $x : T$.

@@

\figref{dot-sub} defines the subtyping judgement $\Gamma \ts S
\sub T$ which states that in environment $\Gamma$ type $S$ is a
subtype of type $T$. Subtyping is regular wrt $\wfe$: if
type $S$ is a subtype of type $T$, then $S$ and $T$ are well-formed
and expanding. Though this regularity limits our calculus to
$\wfe$-types, this limitation allows us to show that subtyping is
transitive, as discussed in section~\ref{dot-preservation}.

@@

\subsubsection{Why limit the calculus to $\wfe$-types?}

Currently, the proof of type-safety via logical relations
fundamentally relies on types having an expansion. However, this was
not our original motivation for limiting the calculus to $\wfe$-types.

Originally, subtyping was not regular wrt $\wfe$. Roughly,
all the $\wfe$ preconditions in subtyping were dropped. In this
broader calculus, subtyping transitivity doesn't hold, because of the
rule (\textsc{$\sub$-rfn}) which requires expansion of the left type.

>>>>

%% Relations
% Subtype 
\newcommand{\sub}{<:}
% Type assignment
\newcommand{\typ}{:}
% reduction
\newcommand{\reduces}{\;\rightarrow\;}
% well-formedness
\newcommand{\wf}{\;\mbox{\textbf{wf}}}
\newcommand{\wfe}{\;\mbox{\textbf{wfe}}}
\newcommand{\nswfe}{\mbox{\textbf{wfe}}}

@@

The last rule, (\textsc{new}), assigns types to object creation
expressions. It is the most complex of DOT's typing rules.  To
type-check an object creation $\new y {T_c \ldefs {\seq{l =
      v}\;\seq{m(x) = t}}} t'$, one verifies first that the type $T_c$
is well-formed (see \figref{dot-wf} for a definition of
well-formedness).  One then determines the set of all declarations
that this type carries, using the expansion relation $\expand$ defined
in \figref{dot-two}.  Every type declaration $\Ldecl L S U$ in
this set must be realizable, i.e.\ its lower bound $S$ must be a
subtype of its upper bound $U$.  Every field declaration $\ldecl l V$
in this set must have a corresponding initializing value of $v$ of
type $V$.  These checks are made in an environment which is extended
by the binding $y: T_c$. In particular this allows field values that
recurse on ``self'' by referring to the bound variable $x$. Similarly,
every method declaration $\mdecl m T W$ must have a corresponding
initializing method definition $m(x) = t$. The parameter type $T$ must
be $\nswfe$ (well-formed and expanding; see \figref{dot-wf}), and
the body $t$ must type check to $W$ in an environment extended by the
bindings $y : T_c$ and $x : T$.

@@

\figref{dot-sub} defines the subtyping judgement $\Gamma \ts S
\sub T$ which states that in environment $\Gamma$ type $S$ is a
subtype of type $T$. Subtyping is regular wrt $\nswfe$: if
type $S$ is a subtype of type $T$, then $S$ and $T$ are well-formed
and expanding. Though this regularity limits our calculus to
$\nswfe$-types, this limitation allows us to show that subtyping is
transitive, as discussed in section~\ref{dot-preservation}.

@@

\subsubsection{Why limit the calculus to $\nswfe$-types?}

Currently, the proof of type-safety via logical relations
fundamentally relies on types having an expansion. However, this was
not our original motivation for limiting the calculus to $\nswfe$-types.

Originally, subtyping was not regular wrt $\nswfe$. Roughly,
all the $\nswfe$ preconditions in subtyping were dropped. In this
broader calculus, subtyping transitivity doesn't hold, because of the
rule (\textsc{$\sub$-rfn}) which requires expansion of the left type.
